---
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE)
```

```{r}
library(tidyr)
library(patchwork)
library(dplyr)
library(ggplot2)
library(Seurat)
library(DT)
library(ComplexHeatmap)
library(circlize)
library(tibble)
library(RColorBrewer)


WORK_DIR <- '/home/chenshuting/drug_toxicity/pan_tissue_landscape/'
DATA_DIR <- '/home/chenshuting/drug_toxicity/pan_tissue/'

source('/home/chenshuting/drug_toxicity/Normal_DSS/scripts/plot/visualization.R')


```

# Non-immune

## Read data

Read main table
```{r}
## Get 295 drug information
drug_info <- readRDS('/home/chenshuting/drug_toxicity/prior_knowledge/drug_info_df.rds')

# Drug-tissue-toxicity metric dataframe
drug_tissuetox_df <- readRDS(paste0(WORK_DIR,'Rdata/drug_tissuetox_df/drug_tissuetox_df.rds')) # Generated by 2.3_drug_tissue_list.R, converted by 1.summrize_result_drug_level.R
drug_tissuetox_df <- drug_tissuetox_df %>%
  mutate(
    drug_name = sub("^[^_]+_", "", drug_id_name)
  )

# Sample DSE information
sample_tox_df <- readRDS(paste0(WORK_DIR,'Rdata/sample_tox_df/sample_tox_df.rds'))
sample_tox_df <- sample_tox_df %>%
  mutate(
    drug_name = sub("^[^_]+_", "", drug_id_name)
  )

## Read sample information
sample_info_tissue <- read.csv(paste0(WORK_DIR,'Tables/sample_info_tissue_scPharm.csv'))
```

Split by tissue into list
```{r}
tissues <- unique(drug_tissuetox_df$tissue)

drug_tissuetox_list <- split(drug_tissuetox_df, drug_tissuetox_df$tissue)
sample_tox_list     <- split(sample_tox_df,     sample_tox_df$tissue)
sample_info_list    <- split(sample_info_tissue, sample_info_tissue$Tissue)

```


Top 5 drugs per tissue

```{r}
top5drug_list <- lapply(
  drug_tissuetox_list,
  function(df) {
    df %>%
  mutate(rra_log10 = -log10(scPharm_RRA_score)) %>%
  filter(is.finite(rra_log10)) %>%          # Prevent Inf / NA
  arrange(desc(rra_log10)) %>%
  distinct(drug_id_name, .keep_all = TRUE) %>%  # Keep only most toxic row per drug
  slice_head(n = 5) %>%
  pull(drug_id_name)
  }
)

top5drug_list

```

## Save ST_nonimmune

```{r}
supp_table1 <- drug_tissuetox_df %>%
  mutate(
    drug_id = sub("_.*$", "", drug_id_name),
    cell_type = case_when(
      tissue %in% c("Blood", "Brain") ~ tissue,
      TRUE ~ "Epithelial"
    ),
    RRA_score = scPharm_RRA_score,
    Rank = scPharm_RRA_rank
  ) %>%
  select(drug_id, drug_name, pathway_name, cell_type, tissue, RRA_score, Rank)

output_dir <- paste0(WORK_DIR, "Tables/supplementary_table_drugrank/")
dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)

saveRDS(supp_table1, paste0(output_dir, "supplementary_table_drugrank_nonimmune.rds"))
write.csv(supp_table1, paste0(output_dir, "supplementary_table_drugrank_nonimmune.csv"), row.names = FALSE)

head(supp_table1)
```



## Epithelial{.tabset}


### F2A-C Ridge plots

```{r}
seleted_tissues <- c("Bowel","Skin","Kidney")

plot_list <- list()

for (tissue in seleted_tissues){
  
  drug_id_names <- top5drug_list[[tissue]]
  
  sample_tox_df_tissue <- sample_tox_list[[tissue]]
  sample_tox_df_subset <- sample_tox_df_tissue[
    sample_tox_df_tissue$drug_id_name %in% drug_id_names,
  ]
  
  sample_tox_df_tissue$drug_name <- "All drugs"
  
  plot_df <- rbind(sample_tox_df_subset, sample_tox_df_tissue)
  
  p <- plot_density_ridges_group(
    df = plot_df,
    group_var = drug_name,
    value_var = DSE,
    ref_group = "All drugs",
    title = tissue
  )+ labs(x = "DSS")
  
  plot_list[[tissue]] <- p
}
```

```{r,fig.width=18,fig.height=5}
final_plot <- (plot_list[[1]] + plot_list[[2]] + plot_list[[3]]) +
  plot_annotation(tag_levels = list(c("A")))

final_plot
ggsave(paste0(WORK_DIR,"Figures/Figure2_tissue_aggregate/ABC_ridgeplot.pdf"),width = 18,height = 5)
```


### F2D RRA

```{r}
epi_tissues <- setdiff(names(sample_tox_list), c("Blood", "Brain"))

plot_list <- lapply(
  epi_tissues,
  function(tissue) {

    df_tissue    <- sample_tox_list[[tissue]]
    drugs_tissue <- top5drug_list[[tissue]]

    plot_drug_rank_consistency(
      df_tissue,
      drugs_to_plot = drugs_tissue,
      title = tissue
      ) + theme(legend.position = c(0.5, 0), # Position relative to entire plot
    legend.justification = c(0, 0)) # Anchor: 1=right, 0=bottom
  }
)
```



```{r,fig.width=16,fig.height=12}
final_plot <- wrap_plots(plot_list, ncol = 4) +
  plot_annotation(
    title = "Rank consistency of top-5 epithelial cell susceptibility drugs across normal tissues",
    theme = theme(
      plot.title = element_text(
        size = 22,
        face = "bold",
        hjust = 0.5
      )
    )
  )

final_plot
ggsave(paste0(WORK_DIR,"Figures/Figure2_tissue_aggregate/D_RRA.pdf"),width = 16,height = 12)

```

## Blood and Brain {.tabset}


### SF2A Ridge plot - Blood
```{r}
tissue <- 'Blood'

drug_id_names <- top5drug_list[[tissue]]

sample_tox_df_tissue <- sample_tox_list[[tissue]]
sample_tox_df_subset <- sample_tox_df_tissue[
  sample_tox_df_tissue$drug_id_name %in% drug_id_names,
]

sample_tox_df_tissue$drug_name <- "All drugs"

plot_df <- rbind(sample_tox_df_subset, sample_tox_df_tissue)

p3 <- plot_density_ridges_group(
  df = plot_df,
  group_var = drug_name,
  value_var = DSE,
  title = tissue,
  ref_group = "All drugs"
)+ labs(x = "DSS")

p3
```

### SF2B RRA - Blood
```{r}
tissue <- 'Blood'

df_tissue    <- sample_tox_list[[tissue]]
drugs_tissue <- top5drug_list[[tissue]]

p4 <-  plot_drug_rank_consistency(
  df_tissue,
  drugs_to_plot = drugs_tissue,
  title = tissue
) + theme(legend.position = c(0.45, 0.1), # Position relative to entire plot
          legend.justification = c(0, 0)) # Anchor: 1=right, 0=bottom

p4

```



### SF2C Ridge plot - Brain
```{r}
tissue <- 'Brain'

drug_id_names <- top5drug_list[[tissue]]

sample_tox_df_tissue <- sample_tox_list[[tissue]]
sample_tox_df_subset <- sample_tox_df_tissue[
  sample_tox_df_tissue$drug_id_name %in% drug_id_names,
]

sample_tox_df_tissue$drug_name <- "All drugs"

plot_df <- rbind(sample_tox_df_subset, sample_tox_df_tissue)

p1 <- plot_density_ridges_group(
  df = plot_df,
  group_var = drug_name,
  value_var = DSE,
  title = tissue,
  ref_group = "All drugs"
)+ labs(x = "DSS")

p1
```

### SF2D RRA - Brain
```{r}
tissue <- 'Brain'

df_tissue    <- sample_tox_list[[tissue]]
drugs_tissue <- top5drug_list[[tissue]]

p2 <-  plot_drug_rank_consistency(
  df_tissue,
  drugs_to_plot = drugs_tissue,
  title = tissue
) + theme(legend.position = c(0.5, 0.1), # Position relative to entire plot
          legend.justification = c(0, 0)) # Anchor: 1=right, 0=bottom

p2

```

```{r,fig.width=20,fig.height=4}
final_plot <- (p3 + p4 + p1 + p2 ) +
  plot_layout(ncol = 4) +
  plot_annotation(
    tag_levels = list(c("A"))
  )
final_plot
ggsave(paste0(WORK_DIR,"Figures/supplementary_figures/SF2_brain_blood.pdf"),width = 20,height = 4)
```




# Immune

## Read data
```{r}
## Get 295 drug information
drug_info <- readRDS('/home/chenshuting/drug_toxicity/prior_knowledge/drug_info_df.rds')

# Drug-tissue-toxicity metric dataframe
drug_tissuetox_df_imm <- readRDS(paste0(WORK_DIR,'Rdata/drug_tissuetox_df/drug_tissuetox_df_imm.rds'))
drug_tissuetox_df_imm <- drug_tissuetox_df_imm %>%
  mutate(
    drug_name = sub("^[^_]+_", "", drug_id_name)
  )
# Keep id for Selumetinib (appears twice in one tissue)
drug_tissuetox_df_imm$drug_name[drug_tissuetox_df_imm$drug_id_name=='1062_Selumetinib'] <- 'Selumetinib_r1'
drug_tissuetox_df_imm$drug_name[drug_tissuetox_df_imm$drug_id_name=='1736_Selumetinib'] <- 'Selumetinib_r2'

# Sample DSE information
sample_tox_df_imm <- readRDS(paste0(WORK_DIR,'Rdata/sample_tox_df/sample_tox_df_imm.rds'))
sample_tox_df_imm <- left_join(sample_tox_df_imm,drug_info[,c('drug_id_name','pathway_name')],by='drug_id_name')
sample_tox_df_imm <- sample_tox_df_imm %>%
  mutate(
    drug_name = sub("^[^_]+_", "", drug_id_name)
  )
# Keep id for Selumetinib (appears twice in one tissue)
sample_tox_df_imm$drug_name[sample_tox_df_imm$drug_id_name=='1062_Selumetinib'] <- 'Selumetinib_r1'
sample_tox_df_imm$drug_name[sample_tox_df_imm$drug_id_name=='1736_Selumetinib'] <- 'Selumetinib_r2'

## Read sample information
sample_info_imm <- read.csv(paste0(WORK_DIR,'Tables/sample_info_imm_scPharm.csv')) 

```

Split by tissue into list
```{r}
tissues <- unique(drug_tissuetox_df_imm$tissue)

drug_tissuetox_list <- split(drug_tissuetox_df_imm, drug_tissuetox_df_imm$tissue)
sample_tox_list     <- split(sample_tox_df_imm,     sample_tox_df_imm$tissue)
sample_info_list    <- split(sample_info_imm, sample_info_imm$Tissue)

```


Top 5 drugs per tissue

```{r}
top5drug_list <- lapply(
  drug_tissuetox_list,
  function(df) {
    df %>%
  mutate(rra_log10 = -log10(scPharm_RRA_score)) %>%
  filter(is.finite(rra_log10)) %>%          # Prevent Inf / NA
  arrange(desc(rra_log10)) %>%
  distinct(drug_id_name, .keep_all = TRUE) %>%  # Keep only most toxic row per drug
  slice_head(n = 5) %>%
  pull(drug_id_name)
  }
)

top5drug_list

```

## Save ST1_immune

```{r}
supp_table1_imm <- drug_tissuetox_df_imm %>%
  mutate(
    drug_id = sub("_.*$", "", drug_id_name),
    cell_type = "immune",
    RRA_score = scPharm_RRA_score,
    Rank = scPharm_RRA_rank
  ) %>%
  select(drug_id, drug_name, pathway_name, cell_type, tissue, RRA_score, Rank)

output_dir <- paste0(WORK_DIR, "Tables/supplementary_table_drugrank/")
dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)

saveRDS(supp_table1_imm, paste0(output_dir, "supplementary_table_drugrank_immune.rds"))
write.csv(supp_table1_imm, paste0(output_dir, "supplementary_table_drugrank_immune.csv"), row.names = FALSE)

head(supp_table1_imm)
```


## F2E-G Ridge plots


```{r}
seleted_tissues <- c("Breast","Head_and_Neck","Ovary")

plot_list <- list()

for (tissue in seleted_tissues){
  
  drug_id_names <- top5drug_list[[tissue]]
  
  sample_tox_df_tissue <- sample_tox_list[[tissue]]
  sample_tox_df_subset <- sample_tox_df_tissue[
    sample_tox_df_tissue$drug_id_name %in% drug_id_names,
  ]
  
  sample_tox_df_tissue$drug_name <- "All drugs"
  
  plot_df <- rbind(sample_tox_df_subset, sample_tox_df_tissue)
  
  p <- plot_density_ridges_group(
    df = plot_df,
    group_var = drug_name,
    ref_group = "All drugs",
    value_var = DSE,
    title = tissue
  )+ labs(x = "DSS")
  
  plot_list[[tissue]] <- p
}
```

```{r,fig.width=18,fig.height=5}
final_plot <- (plot_list[[1]] + plot_list[[2]] + plot_list[[3]]) +
  plot_annotation(tag_levels = list(c("E","F","G")))

final_plot
ggsave(paste0(WORK_DIR,"Figures/Figure2_tissue_aggregate/I_ridgeplot_imm.pdf"),width = 18,height = 5)
```




## F2H RRA

```{r}
imm_tissues <- names(sample_tox_list)

plot_list <- lapply(
  imm_tissues,
  function(tissue) {

    df_tissue    <- sample_tox_list[[tissue]]
    drugs_tissue <- top5drug_list[[tissue]]

    plot_drug_rank_consistency(
      df_tissue,
      drugs_to_plot = drugs_tissue,
      title = paste0(tissue)
      ) + theme(legend.position = c(0.4, 0), # Position relative to entire plot
    legend.justification = c(0, 0)) # Anchor: 1=right, 0=bottom
  }
)
```



```{r,fig.width=16,fig.height=12}
final_plot <- wrap_plots(plot_list, ncol = 4) +
  plot_annotation(
    title = "Rank consistency of top-5 immune cell susceptibility drugs across normal tissues",
    theme = theme(
      plot.title = element_text(
        size = 26,
        face = "bold",
        hjust = 0.5
      )
    )
  )

final_plot
ggsave(paste0(WORK_DIR,"Figures/Figure2_tissue_aggregate/J_RRA.pdf"),width = 16,height = 12)

```




# Save ST1 - Merge immune and non-immune data
```{r}
supp_table1_nonimmune <- readRDS(paste0(WORK_DIR, "Tables/supplementary_table_drugrank/supplementary_table_drugrank_nonimmune.rds"))
supp_table1_immune <- readRDS(paste0(WORK_DIR, "Tables/supplementary_table_drugrank/supplementary_table_drugrank_immune.rds"))

supp_table1_combined <- rbind(supp_table1_nonimmune, supp_table1_immune)

output_dir <- paste0(WORK_DIR, "Tables/supplementary_table_drugrank/")
write.csv(supp_table1_combined, paste0(output_dir, "supplementary_table_drugrank.csv"), row.names = FALSE)

nrow(supp_table1_combined)
head(supp_table1_combined)
```

